---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# zoteror - Access the Zotero API in R

This package aims to introduce basic functionalities to access the Zotero API. Its main goal at this stage is to have enough functionalities to facilitate resizing the storage space used, by ordering items by attachment size, and by allowing to add items to a collection if certain criteria are met.

It now tenatively allows to create new Zotero items, and to take a csv file (or data frame) and import it into Zotero, as long as data are properly mapped. `zoteror` has function that facilitate giving to tabular data a structure that can properly be read into Zotero. See at the bottom of this document for an example [this is still work-in-progress and not functional for all item/attribute types].

# Install

```{r, echo=TRUE, eval=FALSE}
if(!require("remotes")) install.packages("remotes")
remotes::install_github(repo = "giocomai/zoteror")
```

# What works at this stage

## Create an API key and store credentials
```{r, echo=TRUE, eval=FALSE}
credentials <- zot_auth(cache = TRUE)
```
The verification code that appears at the end of the URL after authorization in browser should be input as verification PIN. If the parameter store is enabled - AuthZot(cache=TRUE) - `zoteror` stores the credentials in a local file called "zotero_credentials.rds", which should be considered confidential since it provides access to a given Zotero account. If a pre-existing "zotero_credentials.rds" exists, it is loaded automatically.

N.B. At this stage, it may be easier to actually login on Zotero, go to the (Feeds/API page in Settings)[https://www.zotero.org/settings/keys], and create an API from there.

### Set options
Instead of inputting `user` and `credentials` each time a function is run, it is easier to set them with `ZotSetOptions` at the beginning of every session:

```{r, echo=TRUE, eval=FALSE}
zot_set_options(user = 12345, credentials = "<API>")
```

The respective parameters can then be left empty (`NULL`, which is default) when calling functions. 

## Create a collection
```{r, echo=TRUE, eval=FALSE}
key <- zot_create_collection(collection_name = "zoteror") 
```
Creates a new collection by the given name, and outputs its key. If a collection with the same name already exists, it does not create a new one, but rather outputs the key of the pre-existing collection. `zot_create_collection()` requires an API key with write access.

## Extract details of an item
```{r, echo=TRUE, eval=FALSE}
item <- zot_read_item(id = "<item_id>")

```
Outputs a list with all available information on the item. 

## In which collection(s) is an item?
`zot_which_collection` allows to find out in which collection is an item. If the ID given refers to a 'child item' (e.g. a pdf attachment to a journal article), the function looks for the collection(s) in which the parent item is included. [attachments have separate IDs, and are thus to be found in a collection only if the 'parent item' is]

```{r, echo=TRUE, eval=FALSE}
item <- zot_wich_collection(id = "<item_id>")
```

## Add an item to a collection

```{r, echo=TRUE, eval=FALSE}
zot_add_to_collection(id = "<item_id>", collection_id = "<collection_id>")
```

## Calculate size of all locally stored zotero items

```{r, echo=TRUE, eval=FALSE}
size <- zot_size(path = "/home/user/.zotero/XXXXXX.default/zotero/storage")
# by default, this was inside .mozilla/firefox subdirectory in earlier versions
```

It requires the full path to the local Zotero folder. Outputs size of stored items in bytes and in human-readable Mb. 

# Use case: add all items with attachments bigger than 5MB to a collection

```{r, echo=TRUE, eval=FALSE}
library("zoteror")
library("dplyr")

zot_set_options(user = 12345, credentials = "<API>") # insert user id and API credentials
size <- zot_size(path = "/home/user/.zotero/XXXXXX.default/zotero/storage") # full path to Zotero storage folder

bigIDs <- size %>%
  filter(Size>5000000) %>%
  pull(ID) # filters items larger than 5MB

big_collection_id <- zot_create_collection(collectionName = "plus5") #creates collection "plus5", and if already existing simply outputs its key

for (i in bigIDs) {
    try(zot_add_to_collection(id = bigIDs$ID[i]i,
                              collection_id = big_collection_id))
}

```
This adds all items larger tan 5Mb to a collection called "plus5". `zot_add_to_collection()` is inside `try` to prevent timeout and other errors to stop the script. While running the script outputs the title of items being added to the collection. Currently it goes quite slowly, thus giving time to stop the script if something odd happens. API would allow much more efficient ways of bulk changing items; when the package will work more efficiently, it will still allow to keep the process artificially slow in order to monitor potential oddities.

# Use case: Import tabular data into Zotero with `zoteror`

In order to make sure your data match Zotero fields for a given item type, you can first create a csv template for the given type and paste your data there. Not all columns need to be filled.

```{r eval=TRUE}
library(zoteror)
zot_create_csv_template(item_type = "book", cache = FALSE) 
```

If you enable `cache=TRUE` it will store a csv file in the `zot_csv_templates` sub-folder of the current working directory.

Translation in other languages of all fields are available with:

```{r}
zot_get_item_types_fields(item_type = "book",
                          cache = "FALSE",
                          locale = "it")
```

Internally, Zotero does not store data in a tabular format, so some transformation will be necessary.

For example, if we have data in this format:

```{r}
europe_books <- 
    tibble::tribble(~itemType, ~creators, ~title, ~tags,
                    "book", "Spinelli, Altiero; Rossi, Ernesto", "Il Manifesto di Ventotene", "europe; history")

europe_books
```

We need first to transform them in a format that fully mirrors Zotero's data structure:

```{r message=FALSE}
library(dplyr)

europe_books_zot <- 
    europe_books %>% 
    mutate(creators = zot_convert_creators_to_df_list(creator = creators), 
           tags = zot_convert_tags_to_df_list(tags = tags))

europe_books_zot

```

Having previously set your credentials, you can now upload it to your Zotero accounts.

```{r eval=FALSE}
zot_create_items(item_df = europe_books_zot,
                 collection = "europe")
```

[full support for all data types and more appropriate output will be added in a future version]
